Toxic Code Snippets on Stack Overflow

### 2017-11-13 cloverflow-tse/cloverflow.tex
Incorporated Rocco's changes.

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples in questions and answers. % They
Due to an absence of a checking mechanism after the code has been copied to
Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
being outdated or violating the original software license. We present a study
incorporating several state-of-the-art tools and techniques to automatically
extract and filter detected online clone pairs between 72,365 Java code snippets
on Stack Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
Our investigation of the online clone pairs revealed strong evidence that 154
clones have been copied from a Qualitas project to Stack Overflow.  We found 101
of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
we found 214 code snippets on Stack Overflow that could potentially violate the license of their
original software. We surveyed 201 high-reputation Stack Overflow answerers
(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
code. Only 3 answerers explicitly include software licenses in their answers and
138 answerers (69\%) never check for licensing conflicts between their copied code
snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

@@

Our survey of Stack Overflow answerers shows that online code cloning occurs on
Stack Overflow. Stack Overflow answerers frequently
clone code from other locations, such as their personal projects, company
projects, and open source projects, to Stack Overflow as a solution or
a complement to a solution. The code cloning activity on Stack Overflow
is obviously beneficial considered the popularity of Stack Overflow and its
influence on software development~\cite{Ponzanelli2013,Ponzanelli2014,Park2014}. On the other hand,
there is also a downside caused by low quality, defective, and harmful code
snippets that are reused without an awareness by millions of users.

@@

Toxic code snippets mean code snippets that are harmful for reuse and, in
several cases, are caused by online code cloning. Stack Overflow code snippets
cloned from open source software or online sources can become toxic when they
are (1)~outdated or (2)~violating their original software
license. Outdated code snippets
can be harmful since they are not up-to-date with their originals and may
contain defects. Code snippets from open source projects usually fall under a
specific software license , e.g.\ GNU General Public License (GPL). If they are
cloned to Stack Overflow answers without the license, and then flow to other projects
with conflicting licenses, legal issues may occur.

We would like to motivate the readers by giving two examples of toxic
cod snippets.
The first example is an outdated and potentially license-violating online code clone in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{Hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part
of the accepted answer. The snippet shows how \textsf{Hadoop} implements the
{\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}}
class. The code snippet on the right shows another version of the same method,
but at this time extracted from the latest version (as of October 3, 2017) of
\textsf{Hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added on November
21, 2014. The added line is intended for cleaning up the reference in the
{\small{\verb|buffer|}} variable and avoid excess heap usage
(issue no.~HADOOP-11323\footnote{https://issues.apache.org/jira/browse/HADOOP-11323}).
While this change has already been introduced into the
{\small{\texttt{compare}}} method several years ago, the code example in Stack
Overflow post is still unchanged. 
%This example shows that inconsistencies
%between online code clones and their originals can lead users of Stack Overflow
%to reuse outdated code. 
In addition, the original code snippet of
{\small\texttt{WritableComparator}} class in Hadoop is distributed with Apache license
version 2.0 while its cloned instance on Stack Overflow contains only the
{\small{\texttt{compare}}} method and ignores its license statement on top of the
file. There are two potential issues for this. First, the code snippet may appear to be 
under Stack Overflow's CC BY-SA 3.0 instead of its original Apache
license. Second, if the code snippet is copied and incorporated into another
software project with a conflicting license, a legal issue may arise.

@@

The two toxic code snippet in our examples have been posted on March 11, 2014 and
April 9, 2009 respectively. They have already been viewed 259 and 2,886 times at
the time of writing this paper (October 3, 2017). Our calculation finds that
there will be a new viewer of the first toxic snippet approximately every 5
days compared to almost every day for the second one. A quick web search shows
that the defective {\small{\texttt{humanReadableInt}}} method in the second
example already appears in approximately 500 projects on GitHub. Considering the
popularity of Stack Overflow, which has more than 50 million developers visiting
each month\footnote{Data as of 21 August 2017 from:
	\url{https://stackoverflow.com}}, one toxic code snippets on Stack Overflow can
spread and grow to hundred or thousand copies within only a year or two.

@@

\textbf{Stack Overflow:} We extracted Java code snippets from a snapshot of a
Stack Overflow dump\footnote{\url{https://archive.org/details/stackexchange}} in
January 2016. % The archived dump has a size of 9 gigabytes.
The data dump is in XML, and it contains information of posts (questions and
answers). We were interested in code snippets embedded in posts which were
located between {\small\texttt{<code>}...\texttt{</code>}} tags. A Stack
Overflow thread contains a question and several answers. An answer can also be
marked as an \textbf{accepted answer} by the questioner if the solution fixes
his/her problem. We collected Java code snippets using two criteria. First, we
only focused on code snippets in accepted answers. We chose the snippets in
accepted answers because they actually solved the problems in the questions.
Moreover, they are usually displayed just below the questions which makes them
more likely to be reused than other answers. Moreover, we found that in our data set
of Java answers, 86\% of the accepted answers are, at the same time, the highest voted
answers. Thus, they are well accepted by the Stack Overflow community.
Second, we were only interested in
code snippets of at least ten lines. Although the minimum clone size of six
lines are usual in clone detection~\cite{Bellon2007,Wang2013,Koschke2006}, we
empirically found that snippets of six lines contain a large number of
boiler-plate code of getters/setters, \texttt{equal} or \texttt{hashCode}
methods, which are not interesting for the study. Each snippet was extracted
from the dump and saved to a file. Moreover, we filtered out irrelevant code
snippets that were part of the accepted answers but were not written in Java by
using regular expressions and manual checking.
Finally we obtained 72,365 Java code snippets containing 1,840,581
lines\footnote{Measured by cloc: \url{https://github.com/AlDanial/cloc}} of Java
source code. The median size of the snippets is 17 lines.

@@

We follow Bellon's original definitions of ok-match~\cite{Bellon2007}, which are based on how much two clone
fragments \textit{CF} are contained in each other:

%\squeezeup
%\begin{displaymath}
%overlap(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1) \cup lines(\textrm{\textit{CF}}_2)|} 
%\end{displaymath}
% \squeezeup
\begin{displaymath}
contained(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1)|}
\end{displaymath}
\noindent%
A clone pair \textit{CP} is formed by two clone fragments
\textit{CF$_1$} and \textit{CF$_2$}, i.e.~\textit{CP} =
(\textit{CF$_1$}, \textit{CF$_2$})
%, and the \textit{good-value} and
and the \textit{ok-value} of two clone pairs is defined as

%\squeezeup
%\begin{align*}
%good(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = min(overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1), \\ overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2))
%\end{align*}
\begin{align*}
ok(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) = min(max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1),~~ \\ contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1)),~
\\ max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2),~~ \\contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2)))
\end{align*}

@@

\textbf{Manual investigation:} 
To mitigate the human error, we deployed two persons in the manual clone
investigation process. The first author, who is a research student working on
clone detection research for three years, and the third author, who is a
software engineering research student and familiar with code clones, took the
role of the investigators performing a manual validation and classification of
the merged clone pairs. The two investigators separately went through each clone
pair candidate, looked at the clones, and decided if they are a true positive or
a false positive and classified them into an appropriate pattern. After the
validation, the results from the two investigators were compared. 
%and 607 (26\%)
%conflicts were discussed and resolved. 
There were 338 (15\%) conflicts between 
true and false clones (QS, SQ, EX, UD, BP, IN vs.~AC).
The investigators looked at each conflicting pair together and
discussed until a consensus was made. 
Another 270 pairs (12\%) were
conflicts in the classification of
the true clone pairs. Among these pairs, 145 conflicts were caused by
one investigator being more careful than the other and being able to find evidence of 
copying while the other could not. Thus, resolving the conflicts lead to a better classification,
 i.e.~from UD to QS or EX.
%There were three conflicts where the two
%investigators could not find a consensus and the third author was involved for
%the final judgement.

@@

\textbf{To answer RQ 5.1, although most of the Stack Overflow answerers are aware that their code
can be outdated. 51.5\% of the answerers were very rarely notified while 35.5\% have never been notified of
outdated code in the answers. After being notified, 19.8\% of them rarely or never fix the outdated code.}

@@

\begin{quotation}
\textit{All code posted by me on
	Stack Overflow should be considered public domain without copyright. For
	countries where public domain is not applicable, I hereby grant everyone the
	right to modify, use and redistribute any code posted by me on Stack Overflow
	for any purpose. It is provided ``as-is'' without warranty of any kind.}
\end{quotation}

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples in questions and answers. % They
Due to an absence of a checking mechanism after the code has been copied to
Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
being outdated or violating the original software license. We present a study
incorporating several state-of-the-art tools and techniques to automatically
extract and filter detected online clone pairs between 72,365 Java code snippets
on Stack Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
Our investigation of the online clone pairs revealed strong evidence that 154
clones have been copied from a Qualitas project to Stack Overflow.  We found 101
of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
we found 214 code snippets on Stack Overflow that could potentially violate the license of their
original software. To better understand the phenomenon, we also surveyed 201 high-reputation Stack Overflow answerers
(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
code. Only 3 answerers explicitly include software licenses in their answers and
138 answerers (69\%) never check for licensing conflicts between their copied code
snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

@@

To have a deeper insight on online code clones, we surveyed 201 high-reputation Stack Overflow answerers. The results of such a survey show that online code cloning occurs on
Stack Overflow. Stack Overflow answerers frequently
clone code from other locations, such as their personal projects, company
projects, and open source projects, to Stack Overflow as a solution or
a complement to a solution. The code cloning activity on Stack Overflow
is obviously beneficial considered the popularity of Stack Overflow and its
influence on software development~\cite{Ponzanelli2013,Ponzanelli2014,Park2014}. On the other hand,
there is also a downside caused by low quality, defective, and harmful code
snippets that are reused without an awareness by millions of users.

@@

Toxic code snippets mean code snippets that are harmful for reuse and, in
several cases, are caused by online code cloning. Stack Overflow code snippets
cloned from open source software or online sources can become toxic when they
are (1)~outdated or (2)~violating their original software
license. Outdated code snippets
can be harmful since they are not up-to-date with their originals and may
contain defects. Code snippets from open source projects usually fall under a
specific software license, e.g.\ GNU General Public License (GPL). If they are
cloned to Stack Overflow answers without the license, and then flow to other projects
with conflicting licenses, legal issues may occur.

We would like to motivate the readers by giving two examples of toxic
code snippets.
The first example is an outdated and potentially license-violating online code clone in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{Hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part
of the accepted answer. The snippet shows how \textsf{Hadoop} implements the
{\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}}
class. The code snippet on the right shows another version of the same method,
but at this time extracted from the latest version (as of October 3, 2017) of
\textsf{Hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added on November
21, 2014. The added line is intended for cleaning up the reference in the
{\small{\verb|buffer|}} variable and avoid excess heap usage
(issue no.~HADOOP-11323\footnote{https://issues.apache.org/jira/browse/HADOOP-11323}).
While this change has already been introduced into the
{\small{\texttt{compare}}} method several years ago, the code example in Stack
Overflow post is still unchanged. 
%This example shows that inconsistencies
%between online code clones and their originals can lead users of Stack Overflow
%to reuse outdated code. 
In addition, the original code snippet of
{\small\texttt{WritableComparator}} class in Hadoop is distributed with Apache license
version 2.0 while its cloned instance on Stack Overflow contains only the
{\small{\texttt{compare}}} method and ignores its license statement on top of the
file. There are two potential issues for this. First, the code snippet may appear to be 
under Stack Overflow's CC BY-SA 3.0 instead of its original Apache
license. Second, if the code snippet is copied and incorporated into another
software project with a conflicting license, a legal issue may arise.

@@

The two toxic code snippets in our examples have been posted on March 11, 2014 and
April 9, 2009 respectively. They have already been viewed 259 and 2,886 times at
the time of writing this paper (October 3, 2017). Our calculation finds that
there will be a new viewer of the first toxic snippet approximately every 5
days compared to almost every day for the second one. A quick web search shows
that the defective {\small{\texttt{humanReadableInt}}} method in the second
example already appears in approximately 500 projects on GitHub. Considering the
popularity of Stack Overflow, which has more than 50 million developers visiting
each month\footnote{Data as of 21 August 2017 from:
	\url{https://stackoverflow.com}}, one toxic code snippet on Stack Overflow can
spread and grow to hundred or thousand copies within only a year or two.

@@

\textbf{Stack Overflow:} We extracted Java code snippets from a snapshot of a
Stack Overflow dump\footnote{\url{https://archive.org/details/stackexchange}} in
January 2016. % The archived dump has a size of 9 gigabytes.
The data dump is in XML, and it contains information of posts (questions and
answers). We were interested in code snippets embedded in posts which were
located between {\small\texttt{<code>}...\texttt{</code>}} tags. A Stack
Overflow thread contains a question and several answers. An answer can also be
marked as an \textbf{accepted answer} by the questioner if the solution fixes
his/her problem. We collected Java code snippets using two criteria. First, we
only focused on code snippets in accepted answers. We chose the snippets in
accepted answers because they actually solved the problems in the questions.
Moreover, they are usually displayed just below the questions which makes them
more likely to be reused than other answers. Second, we were only interested in
code snippets of at least ten lines. Although the minimum clone size of six
lines are usual in clone detection~\cite{Bellon2007,Wang2013,Koschke2006}, we
empirically found that snippets of six lines contain a large number of
boiler-plate code of getters/setters, \texttt{equal} or \texttt{hashCode}
methods, which are not interesting for the study. Each snippet was extracted
from the dump and saved to a file. Moreover, we filtered out irrelevant code
snippets that were part of the accepted answers but were not written in Java by
using regular expressions and manual checking.
Finally we obtained 72,365 Java code snippets containing 1,840,581
lines\footnote{Measured by cloc: \url{https://github.com/AlDanial/cloc}} of Java
source code. The median size of the snippets is 17 lines.

@@

We follow Bellon's original definitions of ok-match~\cite{Bellon2007}, which are based on how much two clone
fragments \textit{CF} are contained in each other:
%\squeezeup
%\begin{displaymath}
%overlap(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1) \cup lines(\textrm{\textit{CF}}_2)|} 
%\end{displaymath}
% \squeezeup
\begin{displaymath}
contained(\textrm{\textit{CF}}_1, \textrm{\textit{CF}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1)|}
\end{displaymath}
\noindent%
A clone pair \textit{CP} is formed by two clone fragments
\textit{CF$_1$} and \textit{CF$_2$}, i.e.~\textit{CP} =
(\textit{CF$_1$}, \textit{CF$_2$})
%, and the \textit{good-value} and
and the \textit{ok-value} of two clone pairs is defined as
%\squeezeup
%\begin{align*}
%good(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = min(overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1), \\ overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2))
%\end{align*}
\begin{align*}
ok(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) = min(max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1),~~ \\ contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1)),~
\\ max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2),~~ \\contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2)))
\end{align*}

@@

\textbf{Manual investigation:} 
To mitigate the human error, we deployed two people in the manual clone
investigation process. The first author, who is a research student working on
clone detection research for three years, and the third author, who is a
software engineering research student and familiar with code clones, took the
role of the investigators performing a manual validation and classification of
the merged clone pairs. The two investigators separately went through each clone
pair candidate, looked at the clones, and decided if they are a true positive or
a false positive and classified them into an appropriate pattern. After the
validation, the results from the two investigators were compared. 
%and 607 (26\%)
%conflicts were discussed and resolved. 
There were 338 (15\%) conflicts between 
true and false clones (QS, SQ, EX, UD, BP, IN vs.~AC).
The investigators looked at each conflicting pair together and
discussed until a consensus was made. 
Another 270 pairs (12\%) were
conflicts in the classification of
the true clone pairs. Among these pairs, 145 conflicts were caused by
one investigator being more careful than the other and being able to find evidence of 
copying while the other could not. Thus, resolving the conflicts lead to a better classification,
 i.e.~from UD to QS or EX.
%There were three conflicts where the two
%investigators could not find a consensus and the third author was involved for
%the final judgement.

@@

\textbf{To answer RQ 5.1, although most of the Stack Overflow answerers are aware that their code
can be outdated, 51.5\% of the answerers were very rarely notified and 35.5\% have never been notified of
outdated code in the answers. After being notified, 19.8\% of them rarely or never fix the outdated code.}

@@

\begin{quotation}
\noindent \textit{All code posted by me on
	Stack Overflow should be considered public domain without copyright. For
	countries where public domain is not applicable, I hereby grant everyone the
	right to modify, use and redistribute any code posted by me on Stack Overflow
	for any purpose. It is provided ``as-is'' without warranty of any kind.}
\end{quotation}
